### 2019-08-22 ICES poster/ICES poster.Rmd
poster creation file

---
title: "**Short title:** Flexdashboard poster template"
output: 
  postr::flex_dashboard_poster:
    render_print: false
---

## Header row {data-height=15}

###

<table style="width:100%">
<tr>
  <td><center><h3>
  **Poster title**
  </h3></center></td>
  <td rowspan="2">
  <img src="resources/logo.png" style="width:128px"></img>
  </td>
</tr>
<tr><td><center>
  **Author**
  </center></td>
</tr>
</table>

<hr>

**Abstract** This paragraph will be rendered inside a wide box right below the
poster title and authors information.
The `render_print` format option is set to `TRUE` by default in this poster
sample, and will result in `postr::render` being called when knitting the
document (see `help("flex_dashboard_poster", package = postr)`).
To disable this (recommended in the initial stages of poster development)
set the `render_print` option to `FALSE`.

## Body row {data-height=75}

### **Column 1 title**

#### Heading

This section defines the first column of the body row of the poster.
You can include regular markdown text and R `code chunks`.

```{r}
hist(volcano)
```

### **Column 2 title**

Content for the second column.

To show the code in the chunks you have to ask for it with `echo=TRUE`.

```{r volcano-heat, echo=TRUE, eval=FALSE}
filled.contour(volcano, color.palette = terrain.colors, asp = 1)
```

You can also include data frames as tables just by printing them

```{r}
head(iris)
```

### **Column 3 title**

This template shows 3 columns in the body row, you can change this by adding
or removing third level headings (i.e. lines starting with ###).

## Footer row {data-height=10}

### {data-width=50}

### {data-width=50 .small}